The language of generalization: Probability, vagueness, and interaction

### 2018-03-15 paper/comgen-paper.Rmd
finish light edits to intro; tweak discussion on conceptual knowledge

<!-- Certain linguistic expressions encode generalizations. -->
<!-- Language provides multiple, distinct ways of encoding generalizations. -->
The *language of generalization* covers a diverse swath of natural language expressions.^[
  Some writers refer to the all of the *language of generalization* as "generic language".
  In the literature, however, analysis and experiments are often limited to generalizations about categories. 
  We use the more narrow-scoped terms (e.g., generics, habituals, and causals) to highlight the diversity of semantic types being predicated.
]
Generic language conveys generalizations about categories [e.g., "Dogs have four legs"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008] and is the most well-studied case in the language of generalization.
In contrast to statements about concrete individuals (e.g., "Rufus has four legs"), generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dog}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
Simple events (e.g., "John ran yesterday") can be generalized into habitual sentences (e.g., "John runs").
Even events of complex inferential types such as actual causal events (e.g., "My grandpa drank moonshine and it made him blind") can be described in generalization (e.g., "Drinking moonshine makes you go blind").^[
  Conceretely, we are generalizing over the causal relationship being realized in a particular instance, in this casea person.
]

The language of generalization is ubiquitous in everyday conversation and conveys rich meanings [@Orvell2017].
It is highly-prevalent in child-directed speech [@Gelman2008] and its ability to refer to abstractions beyond the present context has implicated it as central to the growth of conceptual knowledge [@Gelman2004].
It is believed every language can express generalization [@Behrens2005; @Carlson1995], and so understanding the language of generalization is a project with a far-reaching implications.

@@

Despite its ability to convey abstract knowledge, its ubiquity in discourse, and its relative morphosyntactic simplicity, the language of generalization displays subtle context-sensitivities that make it difficult to formalize.
"Robins lay eggs" sounds true and "Robins are female" does not. 
Yet in each case, only 50% of the category have the property (i.e., only the females lay eggs). 
"Mosquitos carry malaria" sounds true despite malaria being present in only a tiny fraction of mosquitos. 
Even more perplexing: "Supreme Court Justices have even social security numbers" is thought to be intuitively a bad sentence even if it were the case that on the current bench, nine out of nine justices have even social securtity numbers (i.e., when the sentence "All Supreme Court Justices have even social security numberes" is true).
Similar context-sensitivity can be observed with habituals: Say Mary climbed mountains 3 times last year, the habitual "Mary climbs mountain" seems well-supported. 
Yet, if John went for a run 3 times in the last year, we probably shouldn't say that "John runs".

@@

These observations have led some to conclude that the literal meaning of a generic statement involves more than just the number of members of the category who have the property, referred to as the *prevalence* of the feature in the category.
<!-- (or the probability of a member of the category having the property e.g.,  $P(x \text{ is friendly} \mid x \text{ is a dog})$) is not sufficient to make generalizations ("Dogs are friendly") true or false. -->
Instead, generic statements are thought of semantically as a direct, linguistic manifestation of conceptual relations [@Leslie2008; @Prasada2000; @Prasada2012]. 
<!-- Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012].  -->
<!-- Indeed, generics can be true for a range of prevalence or frequency levels [@Khelmani2012; @Prasada2013]. -->
For example, the statements "Bishops move diagonally" or "The Speaker of the House succeeds the Vice President" are true not because of a tendency on behalf of the category to uphold the property, but rather the existence of a conceptual relationship (e.g.,  what it means, in the game of chess, to be a bishop).
True generics can be supported by different underlying types of category--property relations [e.g., principled vs. statistical connections; @Prasada2007], and thus support qualitatively different inferences [e.g., "Being striped is one aspect of being a tiger" is well supported while "Carrying malaria is one aspect of being a mosquito" is not not, still while "Tigers are striped" and "Mosquitos carry malaria" sound true; @Prasada2013].
This theory has been influential in psychology, because it predicts several qualitative effects. 
Still, the theory is not sufficiently precise to many any quanitative predictions.

Insofar as there is a single class of linguistic expressions that convey generalization, there should be something common to them all: a literal meaning that unifies generic, habitual, and causal language.
In this paper, we propose such a semantic core.
To do so, we draw upon the tools of Bayesian models of cognition [@Tenenbaum2011] and truth-functional semantics [@Montague1973; @Cohen1999] to formalize the context-dependent aspects of language interpretation [@Frank2012] with the core semantic hypothesis: The meaning of an utterance conveying generalization is vague description of subjective, predictive probabilities.
By its instantiation in a formal model, this hypothesis has meaningful conditions under which it can be tested.
It is also the first theory to make quantitative predictions about human understanding of generalization in language.
<!-- With these ideas, we create an endorsement model, that describes what makes generalizations in language true or false. -->
<!-- Our model is the first to make quantitative predictions about human endorsements of generalizations in language. -->

The paper is organized as follows. 
In Section 1, we describe the computational model, which formalizes the hypothesis that the core meaning of a generalization in language is a vague probability and articulates it in a model of language understanding in which relevant background knowledge can be described formally.
Having defined a model of interpretation, we then describe an information-theoretic, communicative linking function from interpretation to judgements about the truth or acceptibility of a sentence.
The model as articulated predicts fine-grained, quantitative effects of background knowledge, which we both measure and manipulate across our three empirical case studies (Sections 2-4): generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
We compare our model to two previously articulated quantitative models of generics as well as a lesioned-version of our model.
In all cases, we find a very strong agreement of our model's predictions to human elicited endorsements, where the simpler models fall short.
We conclude our paper with a discussion of open-questions for this approach, relation to linguistic and philosophical approaches to generics, and a few sketches of worked examples of other phenomenology in generics.

@@

Generalizations are used to make predictions about events or properties of instances events that an agent has yet to experience [@HumeTHN].
People readily predict that the next dog will have four legs, drinking another cup of coffee will cause jitters, and a new day will find people that we know doing what they habitually do. 
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have feature $f$.
Probability is the universal currency of belief and this prediction can be described by a conditional probability: $P(f \mid k) = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$ (formally, an instance of $k$ will be in the set of things that have $f$).
which we will refer to as the *prevalence* and provide the notational simplification $p = p(f \mid k)$.
Our *prevalence* $p$ is a *prevalence in the mind*, the latent sense that a future instance of a category would have a particular property; others might call this projectibility [@Goodman1955].
The targets of our predictions can vary widely: They may be objects (a dog), events (a coffee drink), or more fine-grained types (a person on a particular day).
The properties also vary (e.g., having fur, causing jitters, going to the gym).
Yet the mathematical description of the inductive belief is always given by a probability $p(f \mid k)$.

<!-- a useful representation  -->
<!-- How do we come to know things?  -->
<!-- How does one acquire inductive beliefs? -->
Probability is a useful representation for human generalization from observations [@Shepard1987].
If you *observe* wugs (a novel category) that have two legs, you might infer that all wugs have two legs.
But not all properties have such strong projectibility (prevalence).
Seeing a wug with broken wings tells you comparatively less about other wugs having broken wings [@Nisbett1983].
Abstract, potentially domain-specific, beliefs about the projectibility of different properties can be represented by a probability distribution over the prevalence $P(p)$ [@Kemp2008].^[
  We adopt the following notational simplification. 
]
Bayes Rule provides the mathematically correct way to update one's prior beliefs with some obesrvational data $d$: $P(p \mid d) \propto P(d \mid p) \cdot P(p)$, assuming some generative process that resulted in the observed data---the likelihood $P(d \mid p)$ [@Tenenbaum2001].

@@

The language of generalization is easy to spot when a property, that could project onto an individual, is predicated of a category (e.g., "Dogs have four legs"; *has four legs* could apply to an individual as in "This dog has four legs").
In the semantics literature on generics, bare plural sentences of this kind are sometimes described as *characterizing sentences* in contrast with *kind-denoting* sentences where the property can only meaningfully apply to the category as a whole (e.g., "Dinosaurs are extinct"; *extinct* cannot apply to an individual).
<!-- Dinosaurs are exctinct ~ John is a smoker? -->
Generalization can also manifest when describing an instances of an individual [*habitual language*; e.g., "John smokes"; @Carlson1977, @Carlson2008].
Verbs like *causes* or *makes* also seem to convey generalization, in this case, about an instance of an actual causal event (*causal language*; e.g., "Staring at the sun makes you go blind"). 
Psychologists, linguists, and philosophers have long studied the language of generalization, as it appears very simple (e.g., syntactically) yet its meaning is difficult to formalize.^[
  We further distinguish the problem of formalizing a *meaning* for the language of generalization as opposed to the problem of *identifying* generalizations. 
  Syntax alone is neither necessary nor sufficient for a listener to know that the sentence conveys a generalization  (e.g., indefinite singulars can encode generalizations: "A dog has four legs" and bare plurals may not: "Dogs are on my front lawn").
  Our analysis thus begins once a sentence has been disambiguated as conveying a generalization. 
]

If generalization from observations can be described by a probability $p$, it is natural to posit that same construct will be at the heart of a semantic theory of the language of generalization.
The semantic tools of *truth values* provide a way to update a state of knowledge by relating the semantic quantity (probability) to  Boolean values [@Montague1973].
The simplest way to relate a scalar quantity like probability is via a *threshold semantics*: The utterance is true if the relevant scalar value is above a threshold.
Thus, the simplest semantics for a generalization about category $k$ with property $f$ is a threshold on the *prevalence*: $\denote{generalization}(p, \theta) = \{p >\theta\}$.

Attempts of this kind have been made before in formal semantics [e.g., by positing that $\theta = 0.5$; @Cohen1999], but heretofore no model has made precise, quantitative predictions about adult judgments. 
In addition to articulating an explicit semantic hypothesis, we present a Bayesian cognitive model that uses this semantics in order to update its beliefs from language.
We then ask: When would a rational speaker choose to produce a generalization (vs. staying silent) to update the Bayesian model's state of knowledge?

@@

The simplest semantics for a generalization is a threshold function: $\denote{gen}(p, \theta) = \{p >\theta\}$, but what threshold is appropriate?
Quantifier statements can be described using particular values of $\theta$: $\denote{some}(p) = \{p > 0\}$; $\denote{all} = \{p = 1\}$.
The extreme flexibility of generalizations (e.g., "Mosquitos carry malaria"; "Birds lay eggs" vs. "Birds are female") suggests that no fixed threshold would suffice. 
Rather than throw out the threshold-semantics, we posit this semantics is *underspecified*.
We hypothesize that $\theta$ is determined by context in a way analogous to how gradable adjectives like "tall" have contextually-determined thresholds [e.g., what counts as "tall" for a four-year-old is different than what counts as "tall" for an adult; @Kennedy2007; @Lassiter2013].

@@

We formalize this underspecified semantics in a Bayesian cognitive model, where the threshold $\theta$ is sampled from a uniform prior distribution over possible thresholds $P(\theta)$ [@Lassiter2013; @Lassiter2015; cf., @Qing2014].
The model also has prior knowledge about the property $P(p)$ and literal interpretation is given by updating these prior beliefs according the truth-functional meaning of utterance $u$:

@@

Following @Lassiter2015, we defined $P(\theta)$ to be a uniform distribution over the support of $P(p)$.

@@

The endorsement model simulates what the interpretation model would believe upon hearing an utterance $u$: $Lit(p, \theta \mid u)$.
$S$ does not know $\theta$ and integrates over the interpreter's posterior beliefs about $\theta$.
Following Bayesian decision theory, we assume the decision to endorse is approximately rational, with degree of rationality given by $\lambda$.
The proportionality in Eq. \ref{eq:S1} implies normalization over a set of alternative utterances. 
We interpret the endorsement task (e.g., "true" vs "false"; "agree" vs. "disagree") as supplying two alternatives: *endorse* or *not*, where *not* we take to be a silent utterance that conveys no information content [@Franke2014cogsci; @Degen2014]: $\denote{null}(x, \theta) = \text{True}$.^[
  This "null" alternative to the generalization ("Ks F") can be realized in at least two other ways: the negation of the generalization (i.e., "It is not the case that Ks F") or the negative generalization (i.e., "Ks do not F").
  All results reported are similar for these two alternatives, and we use the alternative of the "silent" utterance for simplicity.
]

Our endorsement model (Eq. \ref{eq:S1}) provides a mapping from a *referent-category prevalence* $p$  (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) to an endorsement probability for the corresponding generalization (e.g., "Robins lay eggs"): $S(u = u_{gen} \mid p)$. 
The model assumes a threshold semantics based on probability (prevalence) but does not require specifying the threshold *a priori* (note that there is no $\theta$ in the left-hand side of Eq. \ref{eq:S1}).
This model can be seen as a special case of a Rational Speech Act model [@Frank2012; @Goodman2016] with only a single level of recursion and the space of possible utterances maximally constrained; the endorsement model is a "speaker", while the interpretation model is a "listener".

@@

Beliefs about properties result from internal conceptual models of the world. 
Without appealing to domain specific knowledge, one can construct a prior distributions over prevalence by considering the prevalence of the feature among *different* categories.
For example, many different kinds of animals *do not bark* (more precisely: have close to 0 exemplars who bark); thus, the prevalence prior for the feature *barks* $P(p_{barks})$ should have substantial probability mass near 0, because one would expect a randomly-sampled category to have 0 or close to 0 exemplars that bark.
Furthermore, among the categories that do have exemplars which bark (e.g., *dogs*, *seals*, ...), the probability is quite high that an exemplar will bark; thus, the corresponding prevalence prior would have some probability mass at relatively high prevalences, potentially with some uncertainty (Figure\ \@ref(fig:simulations); column 1).
Abstract beliefs between different kinds of categories could manifest in these prevalence priors in the form of multi-modal distributions; for example, the prevalence prior for *lays eggs* should similarly have substantial probability mass near 0 (since many animal categories do not have egg-layers) but should have a secondary component peaked around 0.5, because among the animals with egg-layers, only the female members of the category have the property (Figure\ \@ref(fig:simulations); column 4).

@@

The endorsement model defined in the previous section explicitly predicts that the variability in endorsements for generalizations in language can be explained by two factors: (i) the prevalence communicated in the category under discussion $p$ in Eq. \ref{eq:S1} (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) and (ii) the background knowledge about prevalences $P(p_f)$ (Eq. \ref{eq:L0}).
We simulate how these two factors change both literal interpretation and endorsement using schematic parameter values based on intuition from classic examples in the literature on generic language.
We implemented these and all subsequent Bayesian models in the probabilistic programming language WebPPL [@dippl].
All models, analyses, data and links to experiments used in this paper can be found at \url{https://mhtess.github.io}.

@@

The endorsement model (Eq. \ref{eq:S1}) predicts endorsement probabilities based on the literal interpretation of the utterance and the referent-category prevalence (e.g., how many robins lay eggs; shown in arrows in Figure\ \@ref(fig:simulations)A).
Predictions for a few examples are shown in the bottom row of Figure\ \@ref(fig:simulations).
In accord with intuition, the endorsement model predicts "Dogs bark", "Robins lay eggs", and "Mosquitos carry malaria" are all true, despite the referent-prevalences varying substantially between these three examples. 
Intriguingly, the model predicts "Robins are female" (which has the same prevalence as "Robins lay eggs") should receive intermediate endorsement: It is neither true nor false.
To understand this prediction, examine the interpretation distribution for the *are female* property: It is identical with the prior (Figure\ \@ref(fig:simulations); column 5; top vs. bottom row).
The endorsement model decides that is neither better nor worse to produce the generalization than to stay silent, as they both carry the same information content.
This example can be contrasted with cases such as "Sharks don't eat people" and "Kangaroos have spots", which the model thinks would be misleading to say (i.e., better to say nothing).

These simulations show how the model predicted endorsement depends upon the background knowledge about the property $P(p_f)$ as well as the referent-prevalence $p$ communicated. 
We chose schematic values for each of these, and saw that the endorsement model recovers standard intuitions about the truth or falsity of classic examples from the generics literature.
In what follows, we test this theory for a wide range of generalizations, including generalizations of different types (categories, events, and causes).
We do this by both measuring and manipulating background knowledge and referent-prevalence and relate it to human elicited endorsements of generalizations.

@@

As a formal theory of the language of generalization, our framework unifies large swaths of language about distinct semantic types: "Robins lay eggs" (categories), "Sally smokes" (events), "Herb X makes animals sleepy" (causes).
Generalization underlies each of these, and we formalize their meaning by appealing to prevalence, which we take to mean *predictive probability* (as opposed to frequency).
We test this theory by exploring its implications for *endorsement*, operationalized as a speaker's decision about whether or not to produce the generalization to naive listener. 
Endorsement, we show, depends in systematic ways on the listener's prior beliefs about the prevalence.
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
This framework accounts for a number of philosophically puzzling generic statements (Expt. 1), naturally extends to other domains of generalizations (habituals and causals; Expts. 2 \& 3), and provides an avenue for top-down moderators on language use by way of *predictive* probability (Expt. 2c).
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->

@@

## Prior knowledge and kind representations

>>>>

<!-- Certain linguistic expressions encode generalizations. -->
<!-- Language provides multiple, distinct ways of encoding generalizations. -->
The *language of generalization* covers a diverse swath of natural language expressions.
Generic language conveys generalizations about categories [e.g., "Dogs have four legs"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008] and is the most well-studied case in the language of generalization.^[
  Some writers refer to the all of the *language of generalization* as "generic language".
  In the literature, however, analysis and experiments are often limited to generalizations about categories. 
  We use the more narrow-scoped terms (e.g., generics, habituals, and causals) to highlight the diversity of semantic types being predicated.
]
In contrast to statements about concrete individuals (e.g., "Rufus has four legs"), generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dog}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
Simple events (e.g., "John ran yesterday") can be generalized into habitual sentences (e.g., "John runs"), and even events of complex inferential types such as actual causal events (e.g., "My grandpa drank moonshine and it made him blind") can be described in generalization (e.g., "Drinking moonshine makes you go blind").

Understanding the language of generalization is a project with far-reaching implications.
The language of generalization is ubiquitous in everyday conversation and conveys rich meanings, impacting motivation [@Cimpian2007], transmitting stereotyped beliefs about social categories [@Rhodes2012], making meaning from experience [@Orvell2017].
It is highly-prevalent in child-directed speech [@Gelman2008] and its ability to refer to abstractions beyond the present context suggests its centrality to the growth of conceptual knowledge [@Gelman2004].
Finally, it is believed every language can express generalization [@Behrens2005; @Carlson1995].

@@

Despite its ability to convey abstract knowledge, its ubiquity in discourse, and its relative morphosyntactic simplicity, the language of generalization displays subtle context-sensitivities that make it difficult to formalize.
"Robins lay eggs" sounds true and "Robins are female" does not. 
Yet in each case, only 50% of the category have the property (i.e., only the females lay eggs). 
"Mosquitos carry malaria" sounds true despite malaria being present in only a tiny fraction of mosquitos. 
Even more perplexing: "Supreme Court Justices have even social security numbers" is thought to be intuitively a bad sentence even if it were the case that on the current bench, nine out of nine justices have even social securtity numbers [i.e., when the sentence "All Supreme Court Justices have even social security numbers" is true; @Cohen1999].
Similar context-sensitivity can be observed with habituals: Say Mary climbed mountains 3 times last year, the habitual "Mary climbs mountain" seems well-supported. 
Yet, if John went for a run 3 times in the last year, we probably shouldn't say that "John runs".

@@

These observations have led some to conclude that the literal meaning of a generic statement involves more than just the number of members of the category who have the property, otherwise known as the *prevalence* of the feature in the category.
<!-- (or the probability of a member of the category having the property e.g.,  $P(x \text{ is friendly} \mid x \text{ is a dog})$) is not sufficient to make generalizations ("Dogs are friendly") true or false. -->
Instead, generic statements are thought of semantically as a direct, linguistic manifestation of conceptual relations [@Leslie2008; @Prasada2000; @Prasada2012]. 
<!-- Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012].  -->
<!-- Indeed, generics can be true for a range of prevalence or frequency levels [@Khelmani2012; @Prasada2013]. -->
For example, the statements "Bishops move diagonally" or "The Speaker of the House succeeds the Vice President" are true not because of a tendency on behalf of the category to uphold the property, but rather the existence of a conceptual relationship (e.g.,  what it means, in the game of chess, to be a bishop).
True generics can be supported by different underlying types of category--property relations [e.g., principled vs. statistical connections; @Prasada2006], and thus support qualitatively different inferences [e.g., "Being striped is one aspect of being a tiger" is endorsed while "Carrying malaria is one aspect of being a mosquito" is not, even while "Tigers are striped" and "Mosquitos carry malaria" sound true; @Prasada2013].
This conceptual view of generics has been influential in psychology, because it predicts qualitative differences between different kinds of generics. 
<!-- Still, the theory is not sufficiently precise to make any quantitative predictions. -->

Insofar as there is a single class of linguistic expressions that convey generalization, there should be something common to them all: a literal meaning that unifies generic, habitual, and causal language.
In this paper, we propose such a semantic core.
To do so, we draw upon the tools of Bayesian models of cognition [@Tenenbaum2011] and truth-functional semantics [@Montague1973; @Cohen1999] to formalize the context-dependent aspects of language interpretation [@Frank2012] jointly with the core semantic hypothesis: The meaning of an utterance conveying generalization is a vague description of subjective, predictive probabilities.
By its instantiation in a formal model, this hypothesis has meaningful conditions under which it can be tested.
It is also the first theory to make quantitative predictions about human understanding of generalization in language.
<!-- With these ideas, we create an endorsement model, that describes what makes generalizations in language true or false. -->
<!-- Our model is the first to make quantitative predictions about human endorsements of generalizations in language. -->

The paper is organized as follows. 
In Section 1, we describe the computational model, which formalizes the hypothesis that the core meaning of a generalization in language is a vague probability and articulates it in a model of language understanding in which relevant background knowledge is formally described.
Having defined a model of interpretation, we then describe an information-theoretic, communicative linking function from interpretation to judgments about the truth or acceptibility of a sentence.
The model as articulated predicts fine-grained, quantitative effects of background knowledge, which we both measure and manipulate across our three empirical case studies (Sections 2-4): generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
We compare our model to two previously articulated quantitative models of generics as well as a lesioned-version of our model.
In all cases, we find a very strong agreement of our model's predictions to human elicited endorsements, where the simpler models fall short.
We conclude our paper with a discussion of open-questions for this approach, relation to linguistic and philosophical approaches to generics, and a few sketches of worked examples of other phenomenology in generics.

@@

Generalizations are used to make predictions about events or properties of instances events that an agent has yet to experience [@HumeTHN].
People readily predict that the next dog will have four legs, drinking another cup of coffee will cause jitters, and a new day will find people that we know doing what they habitually do. 
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have feature $f$.
Probability is the universal currency of belief and this prediction can be described by a conditional probability: $P(f \mid k) = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$ (formally, an instance of $k$ will be in the set of things that have $f$), which we will refer to as the *prevalence* and provide the notational simplification $p = P(f \mid k)$.
Our *prevalence* $p$ is a *prevalence in the mind*, a latent sense that a future instance of a category would have a particular property; others might call this projectibility [@Goodman1955].
The targets of our predictions can vary widely: They may be objects (a dog), events (a coffee drink), or more fine-grained types (a person on a particular day).
The properties also vary (e.g., having fur, causing jitters, going to the gym).
Yet the mathematical description of the inductive belief is always given by a probability $P(f \mid k)$.

<!-- a useful representation  -->
<!-- How do we come to know things?  -->
<!-- How does one acquire inductive beliefs? -->
Probability is a useful representation for human generalization from observations [@Shepard1987].
If you *observe* wugs (a novel category) that have two legs, you might infer that all wugs have two legs.
But not all properties have such strong projectibility (prevalence).
Seeing a wug with broken wings tells you comparatively less about other wugs having broken wings [@Nisbett1983].
Abstract, potentially domain-specific, beliefs about the projectibility of different properties can be represented by a probability distribution over the prevalence $P(p)$ [@Kemp2008].
Bayes Rule provides the mathematically correct way to update one's prior beliefs with some obesrvational data $d$: $P(p \mid d) \propto P(d \mid p) \cdot P(p)$, assuming some generative process that resulted in the observed data---the likelihood $P(d \mid p)$ [@Tenenbaum2001].

@@

The language of generalization is easy to spot when a property, that could project onto an individual, is predicated of a category (e.g., "Dogs have four legs"; *has four legs* could apply to an individual as in "This dog has four legs").
In the semantics literature on generics, bare plural sentences of this kind are sometimes described as *characterizing sentences* in contrast with *kind-denoting* sentences where the property can only meaningfully apply to the category as a whole (e.g., "Dinosaurs are extinct"; *extinct* cannot apply to an individual dinosaur).
<!-- Dinosaurs are exctinct ~ John is a smoker? -->
Generalization can also manifest when describing an instances of an individual [*habitual language*; e.g., "John smokes"; @Carlson1977, @Carlson2005].
Verbs like *causes* or *makes* also seem to convey generalization, in this case, about an instance of an actual causal event (*causal language*; e.g., "Staring at the sun makes you go blind"). 
Psychologists, linguists, and philosophers have long studied the language of generalization, as it appears very simple (e.g., syntactically) yet its meaning is difficult to formalize.^[
  We further distinguish the problem of formalizing a *meaning* for the language of generalization as opposed to the problem of *identifying* generalizations. 
  Syntax alone is neither necessary nor sufficient for a listener to know that the sentence conveys a generalization  (e.g., indefinite singulars can encode generalizations: "A dog has four legs" and bare plurals may not: "Dogs are on my front lawn").
  Our analysis thus begins once a sentence has been disambiguated as conveying a generalization. 
]

If generalization from observations can be described by a probability $p$, it is natural to posit that same construct will be at the heart of a semantic theory of the language of generalization.
The semantic tools of *truth values* provide a way to update a state of knowledge by relating the semantic quantity (probability) to Boolean values [@Montague1973].
The simplest way to relate a scalar quantity like probability is via a *threshold semantics*: The utterance is true if the relevant scalar value is above a threshold.
Thus, the simplest semantics for a generalization is a threshold on the prevalence: $\denote{generalization}(p, \theta) = \{p >\theta\}$.

This formalism builds on a tradition in formal semantics to try to precisely define the conditions under which a sentence is true or false [@Montague1973].
We go beyond previous semantic models of generics [e.g., @Cohen1999] by embedding this semantic hypothesis in a cognitive model in which relevant contextual knowledge can be formalized.
By doing so, we thus provide an account of how context influences generic understanding and are able to make precise, quantitative predictions about human judgments.

<!-- Attempts of this kind have been made before in formal semantics [e.g., by positing that $\theta = 0.5$; @Cohen1999], but heretofore no model has made precise, quantitative predictions about adult judgments.  -->
<!-- In addition to articulating an explicit semantic hypothesis, we present a Bayesian cognitive model that uses this semantics in order to update its beliefs from language. -->
<!-- We then ask: When would a rational speaker choose to produce a generalization (vs. staying silent) to update the Bayesian model's state of knowledge? -->

@@

The simplest semantics for a generalization is a threshold function: $\denote{gen}(p, \theta) = \{p >\theta\}$, but what threshold is appropriate?
Quantifier statements can be described using particular values of $\theta$: $\denote{some}(p) = \{p > 0\}$; $\denote{all} = \{p = 1\}$.
The extreme flexibility of generalizations (e.g., "Mosquitos carry malaria"; "Birds lay eggs" vs. "Birds are female") suggests that no fixed threshold would suffice. 
Rather than throw out the threshold-semantics, we posit that the threshold is *underspecified* in the literal meaning and is contextually-determined in a way analogous to how gradable adjectives like "tall" have contextually-determined thresholds [e.g., what counts as "tall" for a four-year-old is different than what counts as "tall" for an adult; @Kennedy2007; @Lassiter2013].

@@

We formalize this underspecified (or, *vague*) semantics in a Bayesian cognitive model, where the threshold $\theta$ is sampled from a uniform prior distribution over possible thresholds $P(\theta)$ [@Lassiter2013;   @Lassiter2015; cf., @Qing2014].^[
  Following @Lassiter2015, we defined $P(\theta)$ to be a uniform distribution over the support of $P(p)$.
]
The model also has prior knowledge about the property $P(p)$ (described in more detail below) and literal interpretation is given by updating these prior beliefs according the truth-functional meaning of utterance $\delta_{\denote{u}(p, \theta)}$:

@@

@@

The endorsement model simulates what the interpretation model would believe upon hearing an utterance $u$: $Lit(p, \theta \mid u)$.
$S$ does not know $\theta$ and integrates over the interpreter's posterior beliefs about $\theta$.
Following Bayesian decision theory, we assume the decision to endorse is approximately rational, with degree of rationality given by $\lambda$.
The proportionality in Eq. \ref{eq:S1} implies normalization over a set of alternative utterances. 
We interpret the endorsement task (e.g., "true" vs "false"; "agree" vs. "disagree") as supplying two alternatives: *endorse* or *not*, where *not* we take to be a silent utterance that conveys no information content [@Franke2014cogsci; @Degen2014]: $\denote{null}(x, \theta) = \text{True}$.^[
  This "null" alternative to the generalization can be realized in at least two other ways: the negation of the generalization (e.g., "It is not the case that Robins lay eggs") or the negative generalization (i.e., "Robins do not lay eggs").
  The latter is different because it creates a new threshold (analogous to how "short" and "tall" are modeled with different thresholds).
  All results reported are similar for these two alternatives, and we use the alternative of the "silent" utterance for simplicity.
]

Our endorsement model (Eq. \ref{eq:S1}) provides a mapping from the prevalence in the category being referred to (i.e., how many robins lay eggs)---the *referent prevalence* $p$---to an endorsement probability for the corresponding generalization (e.g., "Robins lay eggs"): $S(u = u_{gen} \mid p)$. 
The model assumes a threshold semantics based on probability (prevalence) but does not require specifying the threshold *a priori* (note that there is no $\theta$ in the left-hand side of Eq. \ref{eq:S1}).
This model can be seen as a special case of a Rational Speech Act model [@Frank2012; @Goodman2016] with only a single level of recursion and the space of possible utterances maximally constrained; the endorsement model is a "speaker", while the interpretation model is a "listener".

@@

Beliefs about properties result from internal, conceptual models of the world. 
We do not take up the task of formalizing particular conceptual models, but rather merely try to measure their implications on beliefs about prevalence. 
Without appealing to domain specific knowledge, one can construct a prior distributions over prevalence by considering the prevalence of the feature among alternative categories.
For example, many different kinds of animals *do not bark* (more precisely: have close to 0 exemplars who bark); thus, the prevalence prior for the feature *barks* should have substantial probability mass near 0, because one would expect a randomly-sampled category to have 0 or close to 0 exemplars that bark.
Furthermore, among the categories that do have exemplars which bark (e.g., *dogs*, *seals*, ...), the probability is quite high that an exemplar will bark; thus, the corresponding prevalence prior would have some probability mass at relatively high prevalences, potentially with some uncertainty (Figure\ \@ref(fig:simulations); column 1).
Abstract beliefs between different kinds of categories could manifest in these prevalence priors in the form of multi-modal distributions; for example, the prevalence prior for *lays eggs* should similarly have substantial probability mass near 0 (since many animal categories do not have egg-layers) but should have a secondary component peaked around 0.5, because among the animals with egg-layers, only the female members of the category have the property (Figure\ \@ref(fig:simulations); column 4).

@@

The endorsement model defined in the previous section explicitly predicts that the variability in endorsements for generalizations in language can be explained by two factors: (i) the referent prevalence in the category under discussion $p$ in Eq. \ref{eq:S1} (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) and (ii) the background knowledge about prevalences---the prevalence prior---$P(p)$ (Eq. \ref{eq:L0}).
We simulate how these two factors change both literal interpretation and endorsement using schematic parameter values based on intuition from classic examples in the literature on generic language.
We implemented these and all subsequent Bayesian models in the probabilistic programming language WebPPL [@dippl].
All models, analyses, data and links to experiments used in this paper can be found at \url{https://mhtess.github.io}.

@@

The endorsement model (Eq. \ref{eq:S1}) predicts endorsement probabilities based on the literal interpretation of the utterance and the referent prevalence (e.g., how many robins lay eggs; shown in arrows in Figure\ \@ref(fig:simulations)A).
Predictions for a few examples are shown in the bottom row of Figure\ \@ref(fig:simulations).
In accord with intuition, the endorsement model predicts "Dogs bark", "Robins lay eggs", and "Mosquitos carry malaria" are all true, despite the referent-prevalences varying substantially between these three examples. 
Intriguingly, the model predicts "Robins are female" (which has the same prevalence as "Robins lay eggs") should receive intermediate endorsement: It is neither true nor false.
To understand this prediction, examine the interpretation distribution for the *are female* property: It is identical with the prior (Figure\ \@ref(fig:simulations); column 5; top vs. bottom row).
The endorsement model decides that is neither better nor worse to produce the generalization than to stay silent, as they both carry the same information content.
This example can be contrasted with cases such as "Sharks don't eat people" and "Kangaroos have spots", which the model thinks would be misleading to say (i.e., better to say nothing).

These simulations show how the model predicted endorsement depends upon the background knowledge of the interpreter about the property $P(p)$ as well as the referent-prevalence of the endorser $p$ communicated. 
We chose schematic values for each of these, and saw that the endorsement model recovers standard intuitions about the truth or falsity of classic examples from the generics literature.
In what follows, we test this theory for a wide range of generalizations, including generalizations of different types (categories, events, and causes).
We do this by both measuring and manipulating background knowledge and referent-prevalence and relate it to human elicited endorsements of generalizations.

@@

As a formal theory of the language of generalization, our framework unifies large swaths of language about distinct semantic types: "Robins lay eggs" (categories), "Sally smokes" (events), "Herb X makes animals sleepy" (causes).
Generalization underlies each of these, and we formalize their meaning by appealing to prevalence, which we define as a predictive probability (as opposed to frequency).
We test this theory by exploring its implications for truth judgments, operationalized as a speaker's decision about whether or not to produce the generalization to naive listener. 
Endorsement, we show, depends in systematic ways on the listener's prior beliefs about the prevalence.
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
This framework accounts for a number of philosophically puzzling generic statements (Expt. 1), naturally extends to other domains of generalizations (habituals and causals; Expts. 2 \& 3), and provides an avenue for top-down moderators on language use by way of predictive probability (Expt. 2c).
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->

@@

## Prevalence priors and conceptual knowledge
